Score-based tests of measurement invariance: Use in practice

### 2013-09-12 frontiers/frontiers.Rnw
include sim2 results section.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@256 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\subsection{Method}
Measurement invariance violations could occur in
each of the three parameters from Simulation 1 (factor loading,factor covariance,unique
variance), and they could also occur in the extra, unmodeled loading. 
Sample size and magnitude of measurementinvariance violation were manipulated 
in the same way as they were in Simulation 1.The tested parameters are the same as Simulation
1,too.

\subsection{Results}

%<<sim2>>=
%## seed for replication
%RNGkind(kind = "default", normal.kind = "default")
%set.seed(1163)

%if(file.exists("sim2.rda")) {
%  load("sim2.rda")
%} else {
%  sim2 <- simulation(sim="sim2", nobs=480, parms=c("extra","extra+loading","extra+var","extra+error"))
%  sim2$nlevels <- factor(sim2$nlevels)
%  levels(sim2$nlevels) <- paste("m=",levels(sim2$nlevels),sep="")
%  sim2$nobs <- factor(sim2$nobs)
%  levels(sim2$nobs) <- paste("n=",levels(sim2$nobs),sep="")
%  levels(sim2$test) <- c("ordmax","ordwmax","catdiff")
%  save(sim2, file="sim2.rda")
%}
%sim2$test <- factor(as.character(sim2$test),
%  levels = c("ordmax","ordwmax","catdiff"),
%  labels = c("ordmax","ordwmax","catdiff"))
%@

\section{General Discussion}

>>>>

\subsection{Method}
Measurement invariance violations could occur in
each of the three parameters from Simulation 1 (factor loading,factor covariance,unique
variance), and they could also occur in the extra, unmodeled loading. 
Sample size and magnitude of measurement invariance violation were manipulated 
in the same way as they were in Simulation 1.The tested parameters are the same as Simulation
1,too.

\subsection{Results}
Results are presented by the parameter that violated measurement invariance.
Figure~\ref{fig:sim21res} displays results when the unmodeled loading violates 
measurement invariance. One can generally observe that 
the first loading and error variance parameter showed high measurement invariance 
violation simultaneously, while factor covariance did not demonstrate violation.

Figure~\ref{fig:sim22res} displays results when the loading $\lambda_{11}$ violates
measurement invariance. Even though lacking one loading associated with the first indicator, 
statistic showed enough power to detect the first loading violation specifically. 
Similar pattern of specificity and power were observed when the violation parameter 
was $\sigma_{12}$ (Figure~\ref{fig:sim23res}) and $\epsilon_{11}$ (Figure~\ref{fig:sim24res}).

In summary, the proposed test statistics appear robust to unmodeled loading parameters,so long
as the umodeled parameter does not violate measurment invariance. If the unmodeled parameter 
dose violate measurement invariance, however, then the tests assign this violation to other 
parameters that do not violate measurement invariance. The impacted parameters include the error
variance and other loadings associated with the manifest variables that had the unmodeled 
loading. Thus, prior to testing measurement invariance, it is important to study the extent to
which the hypothesized model includes all parameters of importance.

<<sim2>>=
## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1163)

if(file.exists("sim2.rda")) {
  load("sim2.rda")
} else {
  sim2 <- simulation(sim="sim2", nobs=480, parms=c("extra","extra+loading","extra+var","extra+error"))
  sim2$nlevels <- factor(sim2$nlevels)
  levels(sim2$nlevels) <- paste("m=",levels(sim2$nlevels),sep="")
  sim2$nobs <- factor(sim2$nobs)
  levels(sim2$nobs) <- paste("n=",levels(sim2$nobs),sep="")
  levels(sim2$test) <- c("ordmax","ordwmax","catdiff")
  save(sim2, file="sim2.rda")
}
sim2$test <- factor(as.character(sim2$test),
  levels = c("ordmax","ordwmax","catdiff"),
  labels = c("ordmax","ordwmax","catdiff"))
  levels(sim2$pars) <- c("Loading1","Covariance","Error1","All Loadings","All Errors") 
@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is the lacking loading.}
\label{fig:sim21res}
\setkeys{Gin}{width=\textwidth}
<<sim21res, fig=TRUE, height=7, width=7>>=
sim2.tmp <- sim2[sim2$test %in% c("ordmax","ordwmax","catdiff")&sim2$parms %in% c("extra"),]
sim2.tmp$test <- factor(sim2.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim2.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is the lacking loading.}
\label{fig:sim22res}
\setkeys{Gin}{width=\textwidth}
<<sim22res, fig=TRUE, height=7, width=7>>=
sim2.tmp <- sim2[sim2$test %in% c("ordmax","ordwmax","catdiff")&sim2$parms %in% c("extra+loading"),]
sim2.tmp$test <- factor(sim2.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim2.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is the lacking loading.}
\label{fig:sim23res}
\setkeys{Gin}{width=\textwidth}
<<sim23res, fig=TRUE, height=7, width=7>>=
sim2.tmp <- sim2[sim2$test %in% c("ordmax","ordwmax","catdiff")&sim2$parms %in% c("extra+var"),]
sim2.tmp$test <- factor(sim2.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim2.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

\begin{figure}[H]
\caption{Simulated power curves for the ordered 
      $\max\mathit{LM}$ tests, the double-max test and the unordered $\max\mathit{LM}$ across three
      levels of the ordinal variable $m$, and measurement invariance 
      violations of 0--4 standard errors (scaled by $\sqrt{n}$). The violation parameter is the lacking loading.}
\label{fig:sim24res}
\setkeys{Gin}{width=\textwidth}
<<sim24res, fig=TRUE, height=7, width=7>>=
sim2.tmp <- sim2[sim2$test %in% c("ordmax","ordwmax","catdiff")&sim2$parms %in% c("extra+error"),]
sim2.tmp$test <- factor(sim2.tmp$test)
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2.tmp$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | nlevels+pars, group = ~ test, data = sim2.tmp, 
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey))
@

\end{figure}
\section{General Discussion}
